---
title: "French data (version 1.0)"
author: 
  - name: "Thomas Delcey"
    affiliation: "Université de Bourgogne"
    email: "Thomas.delcey@u-bourgogne.fr"
  - name: "Aurelien Goutsmedt"
    affiliation: "UC Louvain; ICHEC"
    email: "aurelien.goutsmedt[at]uclouvain.be" 
sidebar: true
date: "2025"
date-format: "YYYY"
format: 
  html: 
    self-contained: true
---

## Repository 

- [![](https://zenodo.org/badge/DOI/10.5281/zenodo.14541427.svg)](https://doi.org/10.5281/zenodo.14541427)
- Latest update: 21/01/2025
- Version: 1.0

## Citation 

If you use our data, please cite the following reference:

"**Delcey Thomas, and Aurélien Goutsmedt. (2024). Becoming an Economist: A Database of French Economics PhDs. Zenodo. https://doi.org/10.5281/zenodo.14541427**" 

```{}

@article{
  title={Becoming an Economist: A Database of French Economics PhDs},
  author={Delcey, Thomas and Goutsmedt, Aurélien},
  journal={Zenodo},
  year={2024},
  doi={https://doi.org/10.5281/zenodo.14541427}
}

```

## Summary table 

For a quick reference or use, you can find the main information about the thesis in the interactive table below. 

::: callout-warning
This is not the complete dataset but rather a summary of the main columns. We strongly encourage you to consult the documentation and download the full data table for a more comprehensive analysis.
:::

```{r}
#| echo: FALSE 
#| warning: FALSE
#| message: FALSE

pacman::p_load(
  "here",
  "tidyverse",
  "tidytext",
  "data.table",
  "DT",
  "glue")


#### PATH ####

#original text path stored in google drive 
if(str_detect(getwd(), "tunex")){
  data_path <- "G:/Mon Drive/Phd_Project/data"
} else {
  if(str_detect(getwd(), "Admin")) {
    data_path <- "G:/.shortcut-targets-by-id/1Lhjzr0rDBjblTPYh9PoVBsi_uupV9z8L/Phd_Project/data"
  } else {
  if(str_detect(getwd(), "thomd")) {
    data_path <- "G:/.shortcut-targets-by-id/1Lhjzr0rDBjblTPYh9PoVBsi_uupV9z8L/Phd_Project/data"
    } else {
      data_path <- "G:/.shortcut-targets-by-id/1Lhjzr0rDBjblTPYh9PoVBsi_uupV9z8L/Phd_Project/data"
    }}
  }

# data type
raw_data_path <- here(data_path, "raw_data")
intermediate_data_path <- here(data_path, "intermediate_data")

# country
FR_raw_data_path <- here(raw_data_path, "FR")
FR_intermediate_data_path <- here(intermediate_data_path, "FR")

# database
FR_thesefr_raw_data_path <- here(FR_raw_data_path, "these_fr")
FR_thesefr_intermediate_data_path <- here(FR_intermediate_data_path, "these_fr")

FR_sudoc_raw_data_path <- here(FR_raw_data_path, "sudoc")
FR_sudoc_intermediate_data_path <- here(FR_intermediate_data_path, "sudoc")

FR_cleaned_data_path <- here(data_path, "cleaned_data/FR")

FR_blog_data_path <- here(FR_intermediate_data_path, "blogs", "blog_gender")

```

```{r}
#| echo: FALSE 
#| warning: FALSE
#| message: FALSE

thesis_metadata <- readRDS(here(FR_cleaned_data_path, "thesis_metadata.rds"))
thesis_edge <- readRDS(here(FR_cleaned_data_path, "thesis_edge.rds"))

authors_supervisor <- thesis_edge %>%
  filter(entity_role %in% c("author", "supervisor")) %>%
  mutate(entity_fullname = paste(entity_firstname, entity_name)) %>%
  group_by(thesis_id) %>%
  summarize(
    author = paste(unique(entity_fullname[entity_role == "author"]), collapse = "; "),
    supervisor = paste(unique(entity_fullname[entity_role == "supervisor"]), collapse = "; ")
  ) %>%
  ungroup()

thesis_metadata %>%
  select(-c(language, language_2, field, accessible, country, duplicates)) %>% 
  left_join(authors_supervisor) %>% 
  DT::datatable(
    extensions = 'Buttons',
    options = list(
      dom = 'Blfrtip',
      buttons = c('excel', 'csv'),
      pageLength = 12
    )
  )

```



