---
title: "French data (version 1.0)"
author: 
  - name: "Thomas Delcey"
    affiliation: "Université de Bourgogne"
    email: "Thomas.delcey@u-bourgogne.fr"
  - name: "Aurelien Goutsmedt"
    affiliation: "UC Louvain; ICHEC"
    email: "aurelien.goutsmedt[at]uclouvain.be" 
sidebar: true
date: last-modified
date-format: long
format: 
  html: 
    self-contained: true
execute:
 echo: false
 warning: false
 message: false
---

## Repository 

- [![](https://zenodo.org/badge/DOI/10.5281/zenodo.14541427.svg)](https://doi.org/10.5281/zenodo.14541427)
- Latest update: 21/01/2025
- Version: 1.0

## Citation 

If you use our data, please cite the following reference:

"**Delcey Thomas, and Aurélien Goutsmedt. (2024). Becoming an Economist: A Database of French Economics PhDs. Zenodo. https://doi.org/10.5281/zenodo.14541427**" 

```{}

@article{
  title={Becoming an Economist: A Database of French Economics PhDs},
  author={Delcey, Thomas and Goutsmedt, Aurélien},
  journal={Zenodo},
  year={2024},
  doi={https://doi.org/10.5281/zenodo.14541427}
}

```

## Summary table 

For a quick reference or use, you can find the main information about the theses in the interactive table below. 

::: callout-warning
This is not the complete dataset but only the main columns of the `thesis_metadata` data frame. We strongly encourage you to consult the documentation and download the full metadata table and the other tables for a more comprehensive analysis.
:::

```{r}
#| echo: FALSE 
#| warning: FALSE
#| message: FALSE
#| output: FALSE

pacman::p_load(
  "here",
  "tidyverse",
  "tidytext",
  "data.table",
  "DT",
  "glue")


#### PATH ####

#original text path stored in google drive 
if(str_detect(getwd(), "tunex")){
  data_path <- file.path(path.expand("~"), "google_drive/Mon Drive/Phd_Project/data")
} else {
  if(str_detect(getwd(), "Admin")) {
    data_path <- "G:/.shortcut-targets-by-id/1Lhjzr0rDBjblTPYh9PoVBsi_uupV9z8L/Phd_Project/data"
  } else {
    if(str_detect(getwd(), "thomd")) {
      data_path <- "G:/.shortcut-targets-by-id/1Lhjzr0rDBjblTPYh9PoVBsi_uupV9z8L/Phd_Project/data"
    } else {
      if(str_detect(getwd(), "agoutsmedt")) {
        data_path <- "/home/agoutsmedt/google_drive/Phd_Project/data"
      } else {
        data_path <- "G:/.shortcut-targets-by-id/1Lhjzr0rDBjblTPYh9PoVBsi_uupV9z8L/Phd_Project/data"
      }}}
  #  print(paste("The path for data is", data_path))
  }

# database
FR_cleaned_data_path <- here(data_path, "cleaned_data/FR")

```

```{r}
#| echo: FALSE 
#| warning: FALSE
#| message: FALSE

thesis_metadata <- readRDS(here(FR_cleaned_data_path, "thesis_metadata.rds"))

thesis_metadata %>%
  select(-c(title_other, abstract_other, language, language_2, field, accessible, country, duplicates)) %>% 
  mutate(across(starts_with("abstract"), ~str_trunc(., 200))) %>%
  DT::datatable(
    extensions = c('Buttons', "KeyTable"),
    options = list(
      dom = 'Blfrtip',
      buttons = c('excel', 'csv'),
      pageLength = 10,
      searchHighlight = TRUE,
      keys = TRUE
      ),
    filter = 'top' # Adds a search bar for each column at the top
  )

```



